---
title: "Models"
output: html_document
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)  # For ggplot, dplyr, and friends
library(broom)      # To convert models into data frames
```

# Cookies

First we'll generate some fake data about cookies and happiness:

```{r make-cookie-data}
cookies_data <- tribble(
  ~happiness, ~cookies,
  0.5, 1,
  2.0, 2,
  1.0, 3,
  2.5, 4,
  3.0, 5,
  1.5, 6,
  2.0, 7,
  2.5, 8,
  2.0, 9,
  3.0, 10,
)
```

```{r plot-cookie-data}
ggplot(cookies_data, aes(x = cookies, y = happiness)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Cookies eaten", y = "Happiness", 
       title = "Relationship between cookies and happiness")
```

```{r build-cookie-model}
cookies_model <- lm(happiness ~ cookies, data = cookies_data)
```

We can see what's in the `cookies_model` object by summarizing it with `summary()`:

```{r summarize-model}
summary(cookies_model)
```

But this is ugly and it's hard to extract the numbers from that text. It's easier to convert the model results into a data frame, or a table that R can work with. `tidy()` will convert all the coefficients into a table; `glance()` will convert all the diagnostics (R², F-statistic, etc.) into a table:

```{r tidy-model-stuff}
tidy(cookies_model, conf.int = TRUE)
glance(cookies_model)
```


# World happiness

```{r}
world_happiness <- read_csv("data/world_happiness.csv")
```


```{r}
model1 <- lm(happiness_score ~ life_expectancy, data = world_happiness)
tidy(model1)
```

```{r}
ggplot(world_happiness, aes(x = life_expectancy, y = happiness_score)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

```{r}
model2 <- lm(happiness_score ~ region, data = world_happiness)

tidy(model2)
```

```{r}
model_life_school <- lm(happiness_score ~ life_expectancy + school_enrollment, 
                        data = world_happiness)

tidy(model_life_school)
```

```{r}
model_life_school_region <- lm(happiness_score ~ life_expectancy + 
                                 school_enrollment + region, 
                               data = world_happiness)

tidy(model_life_school_region)
```
