---
title: "Mobile phone trends"
author: "Andrew Heiss"
date: "January 15, 2020"
output: html_document
---

```{r setup, include=FALSE}
# Make it so all chunks are printed out in the final document
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)  # For manipulating and plotting data
library(googlesheets4)  # For dealing with Google Sheets
```

## Import data 

We can read in the super short survey data that we just collected using the `googlesheets` package. This might take a moment or two to run.

```{r import, echo=FALSE, message=FALSE}
sheets_deauth()  # The spreadsheet is public so we don't need to try to log in

# Load data
responses_raw <- read_sheet("1LrltZDSkTvJGH6URTuUoW8eEBliUXWWlzBPI6gYKbRg")

# Rename some of the long column names
responses <- responses_raw %>% 
  rename(current_phone = `Who made your current mobile phone?`,
         previous_phone = `Who made your previous mobile phone?`,
         next_phone = `When you purchase your next mobile phone, who do you think the manufacturer will be?`)
```

We now have the data in a data frame called `responses` and it has `r nrow(responses)` survey responses in it.

## Tidying the data

Let's look at the first 5 rows of the data to get a sense of how it is formatted.

```{r tidy, echo=FALSE, message=FALSE}
head(responses, 5)
```

This data looks pretty tidy (we'll define what that means later), so we don't have to do anything in this step.

## Data transformation

We'll summarize the results to prepare them for visualization. We'll just count the proportion of Apple responses in each column to get a sense for how Apple ownership has changed in the latest phone upgrade cycle and see how it might change in the future.

```{r transform, echo=FALSE, message=FALSE}
apple_summary <- responses %>% 
  summarize(Past = sum(previous_phone == "Apple")/n(),
            Present = sum(current_phone == "Apple")/n(),
            Future = sum(next_phone == "Apple")/n()
  ) %>% 
  pivot_longer(Past:Future, values_to = "Percent", names_to = "When") %>% 
  mutate(When = factor(When, levels=c("Past", "Present", "Future")))
apple_summary
```

## Visualizing the survey data

We'll visualize the data using a chart that with points to show the proportions of Apple users in each ownership cycle.

```{r visualize, echo=FALSE, message=FALSE}
ggplot(apple_summary, 
       aes(x = When,
           y = Percent)) +
     geom_point() +
     scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
     labs(y = "Percent Apple Owners", x = "") +
  theme_minimal()
```

## Modeling the results

We can understand the overall trend by fitting a linear model to our 3 data points. This is quantitatively meaningless because our times are imprecise, but the linear model gives us a feel for what's going on with Apple device ownership and where it is headed in the future.

```{r model, echo=FALSE, message=FALSE}
ggplot(apple_summary, 
       aes(x = When,
           y = Percent)) +
     geom_point() +
     scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
     labs(y = "Percent Apple Owners", x = "") +
     geom_smooth(aes(x = as.numeric(When)), method = lm, se = FALSE) +
  theme_minimal()
```

