---
title: "Food access and mortality"
author: "YOUR NAME HERE"
date: "DATE GOES HERE"
output: 
  html_document: 
    toc: yes
  pdf_document: 
    latex_engine: xelatex
    toc: yes
  word_document: 
    toc: yes
---

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(huxtable)

# Make sure you look at this dataset by clicking on its name in the Environment
# panel in RStudio. Sort some of the different columns and look around to get a
# feel for what's in the data
food_health <- read_csv("data/food_health_politics.csv") 
```

# Introduction

We're interested in looking at the relationships between food access, mortality, and politics. Do do this, we look at data from three different sources:

- The USDA's [Food Environment Atlas](https://catalog.data.gov/dataset/food-environment-atlas-f4a22)
- The CDC's ["Compressed Mortality File 1999-2015 Series 20 No. 2U, 2016"](http://wonder.cdc.gov/cmf-icd10.html)
- 2016 election results (found all over the internet)

Each row in the dataset is a US county. The main outcome we care about is `mortality_rate`, or the number of deaths per 100,000 people in a county between 2013-2015. Other interesting variables in the dataset include:

- `pct_low_access_pop`: Percent of the county's population with low access to food
- `pct_children_low_access`: Percent of the county's children with low access to food
- `grocery_stores_per_1000`: Number of grocery stores in a county (per 1,000 residents)
- `snap_stores_per_1000`: Number of stores that accept SNAP (food stamps) in a county (per 1,000 residents)
- `fastfood_per_1000`: Number of fast food stores in a county (per 1,000 residents)
- `per_dem_2012`: Percent of the county that voted for Obama in 2012
- `per_dem_2016`: Percent of the county that voted for Clinton in 2016


# Exploratory analysis

## How related are mortality rate and access to food?

```{r cor-mortality-food}
# Notice how this is a little different from what was in the complete example
# with SAT scores. It's not possible to calculate correlations when there is
# missing data. The `use = "complete.obs"` argument here tells R to ignore any
# rows where either mortality_rate or pct_low_access_pop is missing
cor(food_health$mortality_rate, food_health$pct_low_access_pop,
    use = "complete.obs")
```

SAY SOMETHING HERE. This is backwards from what you might expect, since it trends downward (i.e. the mortality rate is lower in counties with a greater proportion of the population with low access to food). Why might that be? Is there really a relationship?

```{r plot-mortality-food, warning=FALSE}
# Use warning=FALSE in the chunk options to remove the warning about missing data
ggplot(food_health, aes(x = pct_low_access_pop, y = mortality_rate)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "% of county with low access to food", 
       y = "Mortality rate (per 100,000 residents)")
```


## How related are mortality rate and the prevalence of fast food restaurants?

```{r cor-mortality-fastfood}

```

SAY SOMETHING HERE

```{r plot-mortality-fastfood}

```


## How related are mortality rate and the prevalence of SNAP stores per 1,000 residents?

```{r cor-mortality-snap}

```

SAY SOMETHING HERE

```{r plot-mortality-snap}

```


## How related are mortality rate and the percent of the county that voted for Democrats in 2016?

```{r cor-mortality-2016}

```

SAY SOMETHING HERE

```{r plot-mortality-2016}

```


# Models

## Does access to food predict mortality?

SAY SOMETHING HERE

```{r model-mortality-food}
model_mortality_food <- lm(mortality_rate ~ pct_low_access_pop,
                           data = food_health)

tidy(model_mortality_food, conf.int = TRUE)
```

```{r model-details-mortality-food}
glance(model_mortality_food)
```


INTERPRET THE COEFFICIENTS AND RESULTS HERE. What happens as the percent of low access to food goes up by 1%? Is that significant? Again, this is backwards from what you'd expect---as the percent of low access goes up, mortality drops. Why might that be? How much do you trust this finding? (Hint: look at the R2 value)

## Do more SNAP stores per person predict mortality?

```{r model-mortality-snap}

```

```{r model-details-mortality-snap}

```


SAY SOMETHING HERE. What happens as the proportion of SNAP stores goes up? Do you trust this number more or less than low access to food?

## Do election results and access to food and SNAP stores predict mortality?

RUN A MODEL THAT PREDICTS MORTALITY WITH A BUNCH OF COVARIATES (i.e. mortality_rate ~ pct_low_access_pop + snap_stores_per_1000 + per_dem_2016 + anything else you want to throw in)

```{r model-mortality-lots-of-things}

```

```{r model-details-mortality-lots-of-things}

```

SAY SOMETHING HERE. Interpret the different coefficients. How predictive is this model (i.e. what's the R2)? Do you believe this model?

## Mortality, contolling for state differences

RUN A MODEL with some number of plausible independent/explanatory variables. Innclude `state` as one of them

```{r model-mortality-state}
# Add other explanatory variables here
model_with_state <- lm(mortality_rate ~ pct_low_access_pop + state,
                           data = food_health)

# This table is 50+ rows long! While it might be interesting to see changes in
# intercept in relation to Alaska (the omitted state here), like how Alabama's
# mortality rate is 137 higher than Alaska's while DC's is 84 lower, it's not
# super helpful. Controlling for state does capture some of the state-specific
# reasons for varying mortality though, so it's good to include. We just don't
# really need to see all those coefficients. To remove them from this table of
# results, filter them out. The "!" in R means "not", so here we're only looking
# at rows that don't start with "state"
tidy(model_with_state, conf.int = TRUE) %>% 
  filter(!str_starts(term, "state"))
```

```{r model-state-mortality-lots-of-things}

```

SAY SOMETHING ABOUT THIS MODEL


# All models at the same time

PUT ALL THE MODEL RESULTS IN THE SAME SIDE-BY-SIDE TABLE HERE

```{r everything-together}
# Right now there's only one model here. Add the others from above (whatever you
# called them) like so: 
# huxreg(model_mortality_food, some_other_model, yet_another_model, etc)

# Also, by default, huxreg will include all the state coefficients, which we
# don't want. We can create a list of all the state coefficient names and tell
# huxreg to omit them.
state_coef_names <- tidy(model_with_state) %>% 
  filter(str_starts(term, "state")) %>% 
  pull(term)

huxreg(model_mortality_food, model_with_state,
       omit_coefs = state_coef_names)
```
