---
title: "Analyzing RCTs"
output: html_document
---

```{r load-libraries-data, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(ggdag)

village_self_selected <- read_csv("data/village_self_selected.csv")
village_randomized <- read_csv("data/village_randomized.csv")
```


## Diagrams

Here we're assuming that pre-program income, age, and sex (1 = male) all confound the relationship between program and post-program income (older, richer, males are more likely to enroll in the program)

```{r}
income_dag <- dagify(post_income ~ program + age + sex + pre_income,
                     program ~ age + sex + pre_income,
                     exposure = "program",
                     outcome = "post_income",
                     labels = c(post_income = "Post income",
                                program = "Program",
                                age = "Age",
                                sex = "Sex",
                                pre_income = "Pre income"),
                     coords = list(x = c(program = 1, post_income = 5, age = 2, sex = 4, pre_income = 3),
                                   y = c(program = 2, post_income = 2, age = 1, sex = 1, pre_income = 3)))

ggdag_status(income_dag, use_labels = "label", text = FALSE) + 
  theme_dag()
```

If we can randomize assignment to the program, we can delete all the arrows going into "Program":

```{r}
# Randomized
income_dag_rct <- dagify(post_income ~ program + age + sex + pre_income,
                         exposure = "program",
                         outcome = "post_income",
                         labels = c(post_income = "Post income",
                                    program = "Program",
                                    age = "Age",
                                    sex = "Sex",
                                    pre_income = "Pre income"),
                         coords = list(x = c(program = 1, post_income = 5, age = 2, sex = 4, pre_income = 3),
                                       y = c(program = 2, post_income = 2, age = 1, sex = 1, pre_income = 3)))

ggdag_status(income_dag_rct, use_labels = "label", text = FALSE) +
  theme_dag()
```


## Finding program effects

When analyzing RCTs, you generally follow these steps:

1. Check that the assignment was balanced (and really random)
2. Calculate the difference in average outcomes across the treatment and control groups

### Actual RCT

How balanced are the covariates?

```{r}
village_randomized %>% 
  group_by(program) %>% 
  summarize(prop_male = mean(sex),
            avg_age = mean(age),
            avg_income = mean(pre_income))
```

What was the difference in outcomes?

```{r}
village_randomized %>% 
  group_by(program) %>% 
  summarize(avg_post_income = mean(post_income))
```

That's 1279.225 - 1179.977, or `r 1279.225 - 1179.977`. 

You can also use regression to find the same thing:

```{r}
model_rct <- lm(post_income ~ program, data = village_randomized)
tidy(model_rct)
```

### Observational data

What if we do this with the observational data instead? How balanced is that group?

```{r}
village_self_selected %>% 
  group_by(program) %>% 
  summarize(prop_male = mean(sex),
            avg_age = mean(age),
            avg_income = mean(pre_income))
```

What was the difference in outcomes?

```{r}
village_self_selected %>% 
  group_by(program) %>% 
  summarize(avg_post_income = mean(post_income))
```

```{r}
model_observational <- lm(post_income ~ program, data = village_self_selected)
tidy(model_observational)
```

What if we close all the backdoors?

```{r}
model_adjusted <- lm(post_income ~ program + age + sex + pre_income, data = village_self_selected)
tidy(model_adjusted)
```
